Backpropagating Gravitational wave events

### 2022-05-04 src/ms.tex
Reorganize result

We apply the method described in section \ref{sec:method} to the first GW event GW150914.
The parameters involved in the analysis and the range we allow for the inference are tabulated in table \ref{tab:parameters}.
The spin of BHs formed in isolated binaries depends strongly on the angular 
momentum transport in massive stars which is still highly uncertain \kb{(see however Bavera+, fuller+)},
therefore we only consider the two component masses in the observable space in the both the 
root-finding stage and the MCMC stage. For progenitor parameters, we characterize each guess with 
five parameters: the two component masses $M_{\rm 1,ZAMS}$ and $M_{\rm 2,ZAMS}$, the orbital period $t_{\rm orb}$, 
the eccentricity $e$, the metallicity $Z$ at ZAMS. Note that the progenitor formation redshift is not 
obtained through the root-finding process because the formation redshift does not intrinsically affect 
the evolution of the binary. Instead, we compute the time it takes for the binary to evolve from 
formation to the merger and add this time back to the lookback time when a GW event is observed and limit the total 
time to merger to by less than $13.7\,\rm{Gyr}$.
Using the total lookback time, we can compute the redshift of the binary at ZAMS formation following the 
\kw{cite Planck 18} cosmological model. For hyper-parameters, we choose parameters that affect 
binary evolution in \texttt{COSMIC} mostly significantly for massive stars, including the common envelope 
efficiency $\alpha$, the fraction of mass accreted onto the accretor during stable mass transfer $f_{\rm acc}$, 
the critical mass ratio which determines whether Roche-overflow mass transfer 
proceeds stably or produces a common envelope when the donor star is on the Hertzsprung gap 
$q_{\rm crit,3}$, and the root mean square of the Maxwellian distribution
employed for CO natal kicks, $\sigma$.

Once we obtain a set of binaries which successfully map the ZAMS parameters and hyper-parameters to BBH merger masses, 
we re-evolve the set of ZAMS parameters with the same physical assumptions as A21, but vary the common envelope efficiency 
to explore how keeping a fixed model which only varies one hyper parameter affects our analysis. 
Figure~\ref{fig:compare_fixed_variable} shows the distribution of ZAMS masses for three models (first through third columns), 
each with a different $\alpha$ but the same hyper-parameters as A21, 
as well as the results of our sampling which allows our hyper-parameters to vary (fourth column). We find that holding the accretion 
efficiency, common envelope and natal kick physics to fixed values greatly reduces the ZAMS parameter space 
that produces GW150914-like mergers. In contrast, by allowing the hyper-parameters to fully span the model uncertainty, we 
find that there are distinct ZAMS parameters which produce GW150914-like mergers and conversely, ZAMS parameters which 
fail to produce GW150914-like mergers \emph{regardless of our hyper-parameter choice.}

\kb{Need to describe the formation channels and gaps here.}

@@

To check the validity of the recovered posterior, it is essential to check whether it is consistent with the original posterior in the observable space.
We reproject the recovered posterior to the observable space, and compare the KL divergence between the original posterior and the reprojected posterior.
For GW150914, the KL divergence between the reprojected posterior and the original posterior is \kw{fill later} on average.
The KL divergence stays stable for reprojecting with different random seeds. 
\kb{Can we show this in the figure somehow? Maybe like KL divergence as a function of reprojection number as a lower small panel?}
\kw{The problem with that is it is basically one dot or a flat line, which is very unentertaining. I wouldn't necessarily want to make a figure out of the KL divergence just because it is pretty boring and distracting.}
In fig.\ref{fig:GW150914_reprojection}, we can see the reprojected posterior agrees very well with the original posterior.
Compared to the result shown in a previous study, our result covers almost all the posterior of GW150914 in the masses space,
which means our result is more consistent with the observation.
\kw{Do we want to draw comparison with Jeff's result?}
\kb{Yes -- but I think we need to point to Jeff's result in the intro as well.}

@@

\begin{figure}
\includegraphics[width=0.49\textwidth]{static/GW150914_m1m2alpha.png}
\caption{The posterior samples of GW150914 in the progenitor masses space, color-coded by the common envelope efficiency $\alpha$ for that sample.}
\label{fig:GW150914_m1_m2_alpha}
\end{figure}

% \kw{Talk about implication on single event posterior.}
Figure \ref{fig:GW150914_m1_m2_alpha} illustrates the importance of inferring both the progenitor parameters and hyper-parameters jointly.
We can see the posterior samples in $M_1-M_2$ space are spatially correlated with $\alpha$,
this means if we fixed the value of $\alpha$ instead of allowing it to jointly vary with all other parameters,
we would miss the samples that do not match the chosen value of $\alpha$.
Also, our method shows it is capable to find mode that is hard to find by previous methods.
\kw{Talk about the high mass branch that is missed by Jeff's paper}

% \kw{Discuss implication in population.}
To illustrate the potential benefit of our method on the population level, we perform the same analysis for all events in GWTC3.
Plotted in figure \ref{fig:GWTC3_alpha_mass} is the constraint on $f_{\rm acc}$ for events passing the KL divergence test sorted by the median of their primary mass in the observable space.
We can see $f_{\rm acc}$ increases as the primary mass of the event increases.
This means the stable mass transfer (SMT) phase of a light binary is preferably non-conservative, whereas the heavier the events are, the more likely their SMT is conservative.

@@

>>>>

We apply the method described in section \ref{sec:method} to the first GW event GW150914.
The parameters involved in the analysis and the range we allow for the inference are tabulated in table \ref{tab:parameters}.
The spin of BHs formed in isolated binaries depends strongly on the angular 
momentum transport in massive stars which is still highly uncertain \kb{(see however Bavera+, fuller+)},
therefore we only consider the two component masses in the observable space in the both the 
root-finding stage and the MCMC stage. For progenitor parameters, we characterize each guess with 
five parameters: the two component masses $M_{\rm 1,ZAMS}$ and $M_{\rm 2,ZAMS}$, the orbital period $t_{\rm orb}$, 
the eccentricity $e$, the metallicity $Z$ at ZAMS. Note that the progenitor formation redshift is not 
obtained through the root-finding process because the formation redshift does not intrinsically affect 
the evolution of the binary. Instead, we compute the time it takes for the binary to evolve from 
formation to the merger and add this time back to the lookback time when a GW event is observed and limit the total 
time to merger to by less than $13.7\,\rm{Gyr}$.
Using the total lookback time, we can compute the redshift of the binary at ZAMS formation following the 
\kw{cite Planck 18} cosmological model. For hyper-parameters, we choose parameters that affect 
binary evolution in \texttt{COSMIC} mostly significantly for massive stars, including the common envelope 
efficiency $\alpha$, the fraction of mass accreted onto the accretor during stable mass transfer $f_{\rm acc}$, 
the critical mass ratio which determines whether Roche-overflow mass transfer 
proceeds stably or produces a common envelope when the donor star is on the Hertzsprung gap 
$q_{\rm crit,3}$, and the root-mean-square of the Maxwellian distribution
employed for CO natal kicks, $\sigma$.

@@

% \kw{Talk about implication on single event posterior.}
Once we obtain a set of binaries which successfully map the ZAMS parameters and hyper-parameters to BBH merger masses, 
we re-evolve the set of ZAMS parameters with the same physical assumptions as A21, but vary the common envelope efficiency 
to explore how keeping a fixed model which only varies one hyper-parameter affects our analysis. 
Figure~\ref{fig:compare_fixed_variable} shows the distribution of ZAMS masses for three models (first through third columns), 
each with a different $\alpha$ but the same hyper-parameters as A21, 
as well as the results of our sampling which allows our hyper-parameters to vary (fourth column). We find that holding the accretion 
efficiency, common envelope and natal kick physics to fixed values greatly reduces the ZAMS parameter space 
that produces GW150914-like mergers. In contrast, by allowing the hyper-parameters to fully span the model uncertainty, we 
find that there are distinct ZAMS parameters which produce GW150914-like mergers and conversely, ZAMS parameters which 
fail to produce GW150914-like mergers \emph{regardless of our hyper-parameter choice.}
\kw{@Katie, can you put a frame around the fourth column highlighting it is this work?}

\kb{Need to describe the formation channels and gaps here.}
\kw{I think this can be very brief. Save the main explanation to the physics paper.}

To check the validity of the recovered posterior, it is essential to check whether it is consistent with the original posterior in the observable space.
We reproject the recovered posterior to the observable space, and compare the KL divergence between the original posterior and the reprojected posterior.
In particular, we compute KL divergence of the samples obtained from root-finding alone and after MCMC sampling.
For GW150914, the KL divergence between the reprojected posterior and the original posterior is \kw{fill later}logits on average if we use the root-finding result alone,
whereas the MCMC result has a higher KL divergence at \kw{fill} logits.
The KL divergence stays stable for reprojecting with different random seeds. 
\kb{Can we show this in the figure somehow? Maybe like KL divergence as a function of reprojection number as a lower small panel?}
\kw{The problem with that is it is basically one dot or a flat line, which is very unentertaining. I wouldn't necessarily want to make a figure out of the KL divergence just because it is pretty boring and distracting.}
In fig.\ref{fig:GW150914_reprojection}, we can see the reprojected posterior from both the root-fidining process alone and the MCMC result agrees well with the original posterior,
even though the agreement between the MCMC result and the original posterior is not perfect. 
This degradation of the agreement comes from the fact that we are using a KDE to estimate the posterior density in the observable space.
Note that the result from root-finding alone and the result from MCMC agrees well in the evolutionary parameter space.
Compared to the result shown in a previous study, our result covers almost all the posterior of GW150914 in the masses space,
which means our result is more consistent with the observation.

@@

To illustrate the potential benefit of our method on the population level, we perform the same analysis for all events in GWTC3.
Plotted in figure \ref{fig:GWTC3_alpha_mass} is the constraint on $f_{\rm acc}$ for events passing the KL divergence test sorted by the median of their primary mass in the observable space.
We can see $f_{\rm acc}$ increases as the primary mass of the event increases.
This means the stable mass transfer (SMT) phase of a light binary is preferably non-conservative, whereas the heavier the events are, the more likely their SMT is conservative.

@@

% Figure \ref{fig:GW150914_m1_m2_alpha} illustrates the importance of inferring both the progenitor parameters and hyper-parameters jointly.
% We can see the posterior samples in $M_1-M_2$ space are spatially correlated with $\alpha$,
% this means if we fixed the value of $\alpha$ instead of allowing it to jointly vary with all other parameters,
% we would miss the samples that do not match the chosen value of $\alpha$.
% Also, our method shows it is capable to find mode that is hard to find by previous methods.
% \kw{Talk about the high mass branch that is missed by Jeff's paper}

% \kw{Discuss implication in population.}
